Treating Similarity with Respect: How to Evaluate Models of Meaning?

### 2016-06-14 paper.tex
Gamification
Updates from ShareLaTeX

To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incomparable pairs. However, that itself requires a priori knowledge of item similarity or belongingness to a category, making the problem circular. To get out of this vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}. But in case of similarity, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies. For example, \textit{traffic} and \textit{water} might be considered to be similar because of a functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.

@@

@@

In contrast to the current evaluation based on correlation, models also need to be tested on geometric arrangement of subordinates around the focal points, as only the proximity based evaluation does not capture this \cite{1986-13502-00119860101}.

\section{Similarity in Context}

In the previous section we focused on the properties of similarity, illustrating it with word-word examples. This section is dedicated to the nuances of similarity measurement between words, phrases and sentences. Here, we assume that similarity requires the agreement of the grammatical roles.

A noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}. The same similarity principle can be applied to phrases as to words. In this case, similarity is measured in \emph{context}, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}. With verbs this idea can be applied to compare transitive verbs with intransitive. For example, \textit{to cycle} is similar to \textit{ride a bicycle}.

@@

\section{Constructions and Language Sub-systems}

Words, sentences and other constructions draw attention to states of affairs around us. This is both pragmatically and semantically speaking the case. As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.  The point here is that some sentences are similar to others with respect to the functions they perform. For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.

Thus, the amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence. \newcite{tomasello1998return} points out that sentence-level constructions show prototype effects similar to those discussed above for lexical systems (e.g.~colours). Consider the following sentences:
\begin{compactitem}
    \item \textit{``John gave Mary the book.''} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
      \item \textit{``John promised Mary the book.''} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
    %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
    %   \item \textit{``John left Mary the book''} is a example of an \emph{Future Transfer} construction. These usually are build around words such as \textit{bequeath, leave, reserve, grant, etc.}
    %   \item \textit{``John allowed Mary one book''} is a example of an \emph{Enabling Conditions for Transfer} construction. These usually are build around words such as \textit{permit, allow, etc.}
    %   \item \textit{``John wrote Mary a book''} is a example of an \emph{Intended Transfer} construction. These usually are build around words such as \textit{bake, build, make, get, grab, win, earn, etc.}
\end{compactitem}
As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what sentences and utterances are similar.

However, these categories work on the semantic-grammatical level, and might be still handled by similarity in context as described in previous section. What about pragmatics? As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world. For example, he points to the fact that if a person asks another person to \textit{``pass the wine''} they are actually asking for the \textit{bottle}. The speaker just attracts attention to an object of perception in a given situation. There are several ways in which that sentence can both be \emph{grammaticalized} and \emph{lexicalized}. Dialogue act tags is another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
%% Something about sentences with the same function in discourse here.

If one conceptualises sentence similarity with respect to a discourse, then one might ask how different sentences fit in to such a discourse. \newcite{Griffiths2015a} tried to construct to versions of the same dialogue using a bottom-up method. They deconstructed a certain dialogue in a given domain---a receptionist scenario---into \textit{greetings}, \textit{directions} and \textit{farewells}. They used a small custom made corpus for this purpose and created the two dialogues by having people rate the individual utterances by friendliness. The resulting two dialogues were surprisingly uneven. The dialogue was supposed to give instructions to a certain location within a building. The ``friendly version'' was very elaborated and consisted of several sentences:
\enumsentence{
The questionnaire is located in room
Q2-102. That is on the second floor. If
you turn to your right and walk down
the hallway. At the end of the floor you
will find the stairs. Just walk up the
stairs to the top floor and go through
the fire door. The room is then straight
ahead.}
The sentence which served the same purpose in the neutral version was a fairly simple sentence:
\enumsentence{
The questionnaire is located in Q2-102.
}
Often the same functions in dialogue can be performed by as little as one word or several phrases or even a complete story.

@@

Also, sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{No} and \textit{I've seen John eating them} are similar sentences if they are answers to the question \textit{Do we have cookies?} Questions the sentences answer are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}. One approach to treat sentences on their own would be to encode the meaning of a sentence into a vector, in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.

% We could also mention \cite{}.

%       How can we measure sentence similarity?
%
%         3) Transformational approach of either Kolmogorov complexity theory or "the number of changes between the referents".
%         4) Truth conditional equivalence and possible words.

>>>>

To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incomparable pairs. However, that itself requires an a priori knowledge of item similarity or belongingness to a category, making the problem circular. To get out of this vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}. But in case of similarity, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies. For example, \textit{traffic} and \textit{water} might be considered to be similar because of a functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.

@@

Given the dynamic nature of score collection, we suggest setting up \emph{a game with a purpose} (see \newcite{VenhuizenBasileEvangBos2013IWCS} for an example) where players are rewarded for contributing their commonality lists.

@@

In contrast to the current evaluation based on correlation, models also need to be tested on the geometric arrangement of subordinates around the focal points, as only the proximity based evaluation does not capture this \cite{1986-13502-00119860101}.

% \section{Similarity in Context}

% In the previous section we focused on the properties of similarity, illustrating it with word-word examples. This section is dedicated to the nuances of similarity measurement between words, phrases and sentences. Here, we assume that similarity requires the agreement of the grammatical roles.

% \section{Constructions and Language Sub-systems}
\section{Sentence Similarity}

To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them. So what are they?

Words, sentences and other constructions draw attention to states of affairs around us. This is both pragmatically and semantically speaking the case. As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.

The point here is that some sentences are similar to others with respect to the \emph{functions} they perform. For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.

Thus, the amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence. \newcite{tomasello1998return} points out that sentence-level constructions show prototype effects similar to those discussed above for lexical systems (e.g.~colours). Consider the following sentences:
\begin{compactitem}
    \item \textit{``John gave Mary the book.''} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
      \item \textit{``John promised Mary the book.''} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
    %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
    %   \item \textit{``John left Mary the book''} is a example of an \emph{Future Transfer} construction. These usually are build around words such as \textit{bequeath, leave, reserve, grant, etc.}
    %   \item \textit{``John allowed Mary one book''} is a example of an \emph{Enabling Conditions for Transfer} construction. These usually are build around words such as \textit{permit, allow, etc.}
    %   \item \textit{``John wrote Mary a book''} is a example of an \emph{Intended Transfer} construction. These usually are build around words such as \textit{bake, build, make, get, grab, win, earn, etc.}
\end{compactitem}

As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what sentences and utterances are similar.

However, these categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}. The same similarity principle can be applied to phrases as to words. In this case, similarity is measured in context, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}. With verbs this idea can be applied to compare transitive verbs with intransitive. For example, \textit{to cycle} is similar to \textit{to ride a bicycle}.

@@

But still, what about pragmatics? As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world. For example, he points to the fact that if a person asks another person to \textit{``pass the wine''} they are actually asking for the \textit{bottle}. The speaker just attracts attention to an object of perception in a given situation.

There are several ways in which that sentence can both be \emph{grammaticalised} and \emph{lexicalised}. For example, \textit{No} and \textit{I've seen John eating them} are similar sentences because they lexicalise the same answer to the question \textit{Do we have cookies?} More generally, this gives rise to dialogue act tags: another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.

Thus, questions, which the sentences answer, are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}. This also motivates the approach of treating sentences on their own and encoding the meaning of a sentence into a vector in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.

If one conceptualises sentence similarity with respect to a discourse, then one might ask how different sentences fit in to such a discourse. \newcite{Griffiths2015a} tried to construct two versions of the same dialogue using a bottom-up method. They deconstructed a certain dialogue in a given domain---a receptionist scenario---into \textit{greetings}, \textit{directions} and \textit{farewells}. They used a small custom made corpus for this purpose and created the two dialogues by having people rate the individual utterances by friendliness. The resulting two dialogues were surprisingly uneven. The dialogue was supposed to give instructions to a certain location within a building. The ``friendly version'' was very elaborated and consisted of several sentences:
\enumsentence{
The questionnaire is located in room
Q2-102. That is on the second floor. If
you turn to your right and walk down
the hallway. At the end of the floor you
will find the stairs. Just walk up the
stairs to the top floor and go through
the fire door. The room is then straight
ahead.}
The sentence which served the same purpose in the neutral version was a fairly simple sentence:
\enumsentence{
The questionnaire is located in Q2-102.
}
Often the same functions in dialogue can be performed by as little as one word or several phrases or even a complete story.

@@
